---
title: "Lotka Volterra models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sir_models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
lang_output <- function(x, lang) {
  cat(c(sprintf("```%s", lang), x, "```"), sep = "\n")
}
cc_output <- function(x) lang_output(x, "cc")
r_output <- function(x) lang_output(x, "r")
plain_output <- function(x) lang_output(x, "plain")
```

# Stochastic Lotka Volterra model definition


$$\begin{align*}
\frac{dR}{dt} &= \alpha R -\beta RC \\
\frac{dC}{dt} &= \delta \beta RC - \gamma C \\
\end{align*}$$

$R$ is the number of prey, $C$ is the number of predators.$\alpha$ is the birth rate of the prey, $\beta$ is the kill rate, $\gamma$ is the death rate of the predators and $\delta$ describes how much a predator benefits from killing prey.

Discretising this model in time steps of width $dt$ gives the following update equations for each time step:

$$\begin{align*}
R_{t+1} &= R_t - n_{RC} + n_R \\
C_{t+1} &= C_t + \delta n_{RC}  - n_C
\end{align*}$$

where
$$\begin{align*}
n_{R} &\sim B(R, 1 - e^{-\alpha \cdot dt}) \\
n_{RC} &\sim B(R, 1 - e^{-\beta C \cdot dt}) \\
n_{C} &\sim B(C, 1 - e^{-\gamma \cdot dt})
\end{align*}$$

## Implementing the SIR model using [`odin.dust`](https://mrc-ide.github.io/odin.dust/)

The above equations can straightforwardly be written out using the odin DSL:

```{r odin_sir, echo = FALSE, results = "asis"}
r_output(readLines(file.path("LotkaVolterraComp2.R")))
```

This is converted to a C++ dust model, and compiled into a library in a single step, using [`odin.dust`](https://mrc-ide.github.io/odin.dust/). Save the above code as a file named `sir.R`. File names must not contain special characters.

```{r}
library(odin.dust)
LotVolComp <- odin.dust::odin_dust("LotkaVolterraComp2.R")
```

### Running the model:

```{r}
LVC_model <- LotVolComp$new(pars = list(dt = 1,
                                     R_ini = 200,
                                     C_ini = 200,
                                     r_R = 0.2,
                                     K1 = 200,
                                     r_C = 0.2,
                                     K2 = 200,
                                     alpha_RC = 0.15,
                                     alpha_CR = 0.1),
                         time = 1,
                         n_particles = 10L,
                         n_threads = 4L,
                         seed = 1L)
```

# initial state: (time, prey, predator)

```{r}
LVC_model$state()
```



# running the model for 10, 20, 30 time steps:

```{r}
LVC_model$run(1)
LVC_model$run(2)
LVC_model$run(3)
```
# okay, all numbers are integers
# but now I simulate twice how the prey and pred interact, i.e. the number of deaths of prey due to it being killed by pred is independent of the increase we see in the pred due to it eating the prey
# that does not make any sense
# instead, I would want the pred grow at a constant * the rate that the prey decreases, I think.

```{r}
dt <- 1
n_particles <- 10L
p_new <- list(dt = 1, R_ini = 200, C_ini = 200, r_R = 0.2, K1 = 200, r_C = 0.2, K2 = 200, alpha_RC = 0.15, alpha_CR = 0.1)
LVC_model$update_state(pars = p_new, time = 0)
LVC_model$state()
```


```{r fig.height=5, fig.width=7}
n_times <- 500
x <- array(NA, dim = c(LVC_model$info()$len, n_particles, n_times))

for (t in seq_len(n_times)) {
  x[ , , t] <- LVC_model$run(t)
}
time <- x[1, 1, ]
x <- x[-1, , ]

par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
cols <- c(R = "#8c8cd9", C = "#cc0044")
matplot(time, (x[1,1 , ]), type = "l",
         xlab = "Time", ylab = "Number of individuals",
         col = cols[["R"]], lty = 1, ylim = range(x))
matlines(time, (x[2,1 , ]), col = cols[["C"]], lty = 1)
legend("left", lwd = 1, col = cols, legend = names(cols), bty = "n")
```

```{r fig.height=5, fig.width=7}
dt <- 1
n_particles <- 10L
p_new <- list(dt = 1, R_ini = 400, C_ini = 400, r_R = 0.2, K1 <- 400, r_C = 0.2, K2 <- 400, alpha_RC = 0.2, alpha_CR = 0.2)
LVC_model$update_state(pars = p_new, time = 0)
LVC_model$state()


n_times <- 500
x <- array(NA, dim = c(LVC_model$info()$len, n_particles, n_times))

for (t in seq_len(n_times)) {
  x[ , , t] <- LVC_model$run(t)
}
time <- x[1, 1, ]
x <- x[-1, , ]

par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
cols <- c(R = "#8c8cd9", C = "#cc0044")
matplot(time, t(x[1, , ]), type = "l",
         xlab = "Time", ylab = "Number of individuals",
         col = cols[["R"]], lty = 1, ylim = range(x))
matlines(time, t(x[2, , ]), col = cols[["C"]], lty = 1)
legend("left", lwd = 1, col = cols, legend = names(cols), bty = "n")
```


### this is the Lotka Volterra model for competition between two species.
# next step is to extend it to three species, then n>3
# also, I should produce "data" with the model and then try to fit the model to it. The question is a bit, which parameters should be fitted...? There are quite a lot of parameters (at least 6 for the 2-species model). With this amount of parameters, I can fit a lot, can't I?



