---
title: "Lotka Volterra models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sir_models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
lang_output <- function(x, lang) {
  cat(c(sprintf("```%s", lang), x, "```"), sep = "\n")
}
cc_output <- function(x) lang_output(x, "cc")
r_output <- function(x) lang_output(x, "r")
plain_output <- function(x) lang_output(x, "plain")
```

# Stochastic Lotka Volterra model definition


$$\begin{align*}
\frac{dR}{dt} &= \alpha R -\beta RC \\
\frac{dC}{dt} &= \delta \beta RC - \gamma C \\
\end{align*}$$

$R$ is the number of prey, $C$ is the number of predators.$\alpha$ is the birth rate of the prey, $\beta$ is the kill rate, $\gamma$ is the death rate of the predators and $\delta$ describes how much a predator benefits from killing prey.

Discretising this model in time steps of width $dt$ gives the following update equations for each time step:

$$\begin{align*}
R_{t+1} &= R_t - n_{RC} + n_R \\
C_{t+1} &= C_t + \delta n_{RC}  - n_C
\end{align*}$$

where
$$\begin{align*}
n_{R} &\sim B(R, 1 - e^{-\alpha \cdot dt}) \\
n_{RC} &\sim B(R, 1 - e^{-\beta C \cdot dt}) \\
n_{C} &\sim B(C, 1 - e^{-\gamma \cdot dt})
\end{align*}$$

## Implementing the SIR model using [`odin.dust`](https://mrc-ide.github.io/odin.dust/)

The above equations can straightforwardly be written out using the odin DSL:

```{r odin_sir, echo = FALSE, results = "asis"}
r_output(readLines(file.path("LotkaVolterraCompN.R")))
```

This is converted to a C++ dust model, and compiled into a library in a single step, using [`odin.dust`](https://mrc-ide.github.io/odin.dust/). Save the above code as a file named `sir.R`. File names must not contain special characters.

```{r}
library(odin.dust)
LotVolCompN <- odin.dust::odin_dust("LotkaVolterraCompN.R")
```
### Define model parameters
```{r}
#params <- list(dt = 1, species_no = 3, C_ini = c(200,200,200), r_C = c(0.2,0.2,0.1), K = c(200,200,200), alpha = matrix(data = c(1,0,.1,.5,1,-.2,.1,.4,1),nrow = 3, ncol = 3, byrow = TRUE))
params <- list(dt = 1, species_no = 4, C_ini = c(200,200,200,200), r_C = c(0.2,0.2,0.1,0.2), K = c(200,200,200,200), alpha = matrix(data = c(1,0,.1,.5,.2,1,.4,-.2,.1,.4,1,.3,.1,-.4,-.2,1),nrow = 4, ncol = 4, byrow = TRUE))
#params <- list(dt = 1, species_no = 9, C_ini = rep(200,9), r_C = rep(.2,9), K = rep(200,9), alpha = diag(9)) #test with nine compartments
```



### Running the model:

```{r}
LVC_modelN <- LotVolCompN$new(pars = params,
                         time = 1,
                         n_particles = 10L,
                         n_threads = 4L,
                         seed = 1L)
```

# initial state: (time, prey, predator)

```{r}
LVC_modelN$state()
```



# running the model for 10, 20, 30 time steps:

```{r}
LVC_modelN$run(1)
LVC_modelN$run(2)
LVC_modelN$run(3)
```

```{r}
dt <- 1
n_particles <- 10L
p_new <- params
LVC_modelN$update_state(pars = p_new, time = 0)
LVC_modelN$state()
```


```{r fig.height=5, fig.width=7}
n_times <- 500
x <- array(NA, dim = c(LVC_modelN$info()$len, n_particles, n_times))

for (t in seq_len(n_times)) {
  x[ , , t] <- LVC_modelN$run(t)
}
time <- x[1, 1, ]
x <- x[-1, , ]

par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
if (length(x[,1,1]) <= 8){
  cols <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") #8 colorblind friendly colors
} else {
  cols <- rainbow(length(x[,1,1]))
}

# probably should find a more dynamic way for models with >8 species?
matplot(time, (x[1,1 , ]), type = "l",
         xlab = "Time", ylab = "Number of individuals",
         col = cols[1], lty = 1, ylim = range(x))
for (species in 2:length(x[,1,1])) {
  matlines(time, (x[species,1 , ]), col = cols[species:length(x[,1,1])], lty = 1)
}

legend("bottomleft", lwd = 1, col = cols[1:length(x[,1,1])], legend = 1:length(x[,1,1]), bty = "n")
```

#### I compared this to the "hand made" 2-species model and the behaviour seems very similar! :-)

### issue with negative values in alpha??? check whether the same issue exists with the poisson distribution or not. hm. Could be that I get a "negative probability"?


```{r fig.height=5, fig.width=7}
dt <- 1
n_particles <- 10L
p_new <- list(dt = 1, C_ini = c(200,200,200), r_C = c(0.2,0.2,0.2), K = c(500,500,500), alpha = matrix(data = c(1,.15,.1,.1,1,.12,.1,.11,1),nrow = 3, ncol = 3, byrow = TRUE, dimnames = list(c("R","C","C2"),c("R","C","C2"))))
LVC_model3v2$update_state(pars = p_new, time = 0)
LVC_model3v2$state()


n_times <- 500
x <- array(NA, dim = c(LVC_model3v2$info()$len, n_particles, n_times))

for (t in seq_len(n_times)) {
  x[ , , t] <- LVC_model$run(t)
}
time <- x[1, 1, ]
x <- x[-1, , ]

par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
cols <- c(R = "#8c8cd9", C = "#cc0044", C2 = "#FFA500")
matplot(time, t(x[1, , ]), type = "l",
         xlab = "Time", ylab = "Number of individuals",
         col = cols[["R"]], lty = 1, ylim = range(x))
matlines(time, t(x[2, , ]), col = cols[["C"]], lty = 1)
matlines(time, (x[3,1 , ]), col = cols[["C2"]], lty = 1)
legend("left", lwd = 1, col = cols, legend = names(cols), bty = "n")
```


### this is the Lotka Volterra model for competition between two species.
# next step is to extend it to three species, then n>3
# also, I should produce "data" with the model and then try to fit the model to it. The question is a bit, which parameters should be fitted...? There are quite a lot of parameters (at least 6 for the 2-species model). With this amount of parameters, I can fit a lot, can't I?



