---
title: "SIR models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sir_models}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
lang_output <- function(x, lang) {
  cat(c(sprintf("```%s", lang), x, "```"), sep = "\n")
}
cc_output <- function(x) lang_output(x, "cc")
r_output <- function(x) lang_output(x, "r")
plain_output <- function(x) lang_output(x, "plain")
```

# Stochastic SIR model definition

A simple definition of the SIR model, as given in the [odin documentation](https://mrc-ide.github.io/odin/articles/discrete.html) is:
$$\begin{align*}
\frac{dS}{dt} &= -\beta \frac{SI}{N} \\
\frac{dI}{dt} &= \beta \frac{SI}{N} - \gamma I \\
\frac{dR}{dt} &= \gamma I \\
\end{align*}$$
$S$ is the number of susceptibles, $I$ is the number of infected and $R$ is the number recovered; the total population size $N = S + I + R$ is constant. $\beta$ is the infection rate, $\gamma$ is the recovery rate.

Discretising this model in time steps of width $dt$ gives the following update equations for each time step:

$$\begin{align*}
S_{t+1} &= S_t - n_{SI} \\
I_{t+1} &= I_t + n_{SI} - n_{IR} \\
R_{t+1} &= R_t + n_{IR}
\end{align*}$$

where
$$\begin{align*}
n_{SI} &\sim B(S, 1 - e^{-\beta \frac{I}{N} \cdot dt}) \\
n_{IR} &\sim B(I, 1 - e^{-\gamma \cdot dt})
\end{align*}$$

## Implementing the SIR model using [`odin.dust`](https://mrc-ide.github.io/odin.dust/)

The above equations can straightforwardly be written out using the odin DSL:

```{r odin_sir, echo = FALSE, results = "asis"}
r_output(readLines(file.path("sir.R")))
```

This is converted to a C++ dust model, and compiled into a library in a single step, using [`odin.dust`](https://mrc-ide.github.io/odin.dust/). Save the above code as a file named `sir.R`. File names must not contain special characters.

```{r}
library(odin.dust)
gen_sir <- odin.dust::odin_dust("sir.R")
```

## Saving a model into a package

If you want to distribute your model in an R package, rather than building it locally, you will want to use the `odin_dust_package()` function instead. This will write the transpiled C++ dust code into `src` and its R interface in `R/dust.R`. Package users are not required to regenerate the dust code, and their compiler will build the library when they install the package. You may also wish to add a file `R/zzz.R` to your package `myrpackage` with the following lines:

```r
##' @useDynLib myrpackage, .registration = TRUE
NULL
```

to help automate the compilation of the model.

# Running the SIR model with [`dust`](https://mrc-ide.github.io/dust/)

Now we can use the `new` method on the generator to make `dust` objects. `dust` can be driven directly from R, and also interfaces with the [`mcstate`](https://mrc-ide.github.io/mcstate/) package to allow parameter inference and forecasting.

[`new()`](https://mrc-ide.github.io/dust/reference/dust.html#method-new) takes the data needed to run the model i.e. a list with any parameters defined as `user` in the odin code above, the value of the initial time step $t_0$, and the number of particles, each for now can simply be thought of as an independent stochastic realisation of the model, but in the next time step will be used when inferring model parameters.

Additional arguments include the number of threads to parallelise the particles over, and the seed for the random number generator. The seed must be an integer, and using the same seed will ensure reproducible results for all particles. To use this to directly create a new dust object with 10 particles, run using 4 threads:

```{r}
sir_model <- gen_sir$new(pars = list(dt = 1,
                                     S_ini = 2000,
                                     I_ini = 5,
                                     beta = 0.15,
                                     gamma = 0.1),
                         time = 1,
                         n_particles = 10L,
                         n_threads = 4L,
                         seed = 1L)
```

The initial state is ten particles wide, four states deep ($t$, $S$, $I$, $R$):

```{r}
sir_model$state()
```

Run the particles (repeats) forward 10 time steps of length $dt$, followed by another 10 time steps:

```{r}
sir_model$run(10)
sir_model$run(20)
```

#```{r}
#case_compare <- function(state, prev_state, observed, pars = NULL) {
#  cases_modelled <- prev_state[1, , ] - state[1, , ] 
#  dpois(incidence_observed, observed$cases, log = TRUE)
# }
#```




We can change the parameters, say by increasing the infection rate and the population size, by reinitalising the model with `reset()`. We will also use a smaller time step to calculate multiple transitions per unit time:

```{r}
dt <- 1
n_particles <- 10L
p_new <- list(dt = dt, S_ini = 2000, I_ini = 10, beta = 0.25, gamma = 0.1)
sir_model$update_state(pars = p_new, time = 0)
sir_model$state()
```

Let's run this epidemic forward, and plot the trajectories:

```{r fig.height=5, fig.width=7}
n_times <- 200
x <- array(NA, dim = c(sir_model$info()$len, n_particles, n_times))

for (t in seq_len(n_times)) {
  x[ , , t] <- sir_model$run(t)
}
time <- x[1, 1, ]
x <- x[-1, , ]

par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
cols <- c(S = "#000000", I = "#E69F00", R = "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") #8 colorblind friendly colors
#cols <- c(S = "#8c8cd9", I = "#cc0044", R = "#999966")
matplot(time, t(x[1, , ]), type = "l",
         xlab = "Time", ylab = "Number of individuals",
         col = cols[["S"]], lty = 1, ylim = range(x))
matlines(time, t(x[2, , ]), col = cols[["I"]], lty = 1)
matlines(time, t(x[3, , ]), col = cols[["R"]], lty = 1)
legend("right", lwd = 1, col = cols, legend = names(cols[1:3]), bty = "n")
```

## Trying to fit the model to "data" (data produced by the model itself)

```{r}
# create a data frame based on the simulation 1 that I can use to fit the model to
df_SIRinfected <- data.frame(unlist(x[2,1,]))
names(df_SIRinfected)[names(df_SIRinfected)=="unlist.x.2..1...."] <- "Infected"
df_SIRinfected$day <- c(1:200)
```

```{r}
#install.packages("drat") # -- if you don't have drat installed
drat:::add("ncov-ic")
#install.packages("mcstate")
library(mcstate)
```

```{r paged.print = TRUE, layout = "l-body-outset"}
#dt <- 0.25
dt <- 1
sir_data <- mcstate::particle_filter_data(data = df_SIRinfected,
                                          time = "day",
                                          rate = 1 / dt)
rmarkdown::paged_table(sir_data)
plot(df_SIRinfected$day, df_SIRinfected$Infected,
     type = "l", xlab = "Day", ylab = "Total cases")
```

```{r}
case_compare <- function(state, observed, pars = NULL) {
  exp_noise <- 1e6
  incidence_modelled <- state[3, , drop = TRUE] # "1," should be time, "2," should be S, "3," should be I, "4," R
  incidence_observed <- observed$Infected
  lambda <- incidence_modelled 
  +
    rexp(n = length(incidence_modelled), rate = exp_noise)
  dpois(x = incidence_observed, lambda = lambda, log = TRUE)
}
```

We can use the `info()` method on a dust model to inspect what order the state variables will come in:

```{r}
gen_sir$new(pars = list(), time = 0, n_particles = 1L)$info()
```
### INCIDENCE FUNCTION weird

```{r}
#incidence_compare <- function(state, observed, pars = NULL) {
#  exp_noise <- 1e6

#  incidence_modelled <- state[3, , drop = TRUE] # "1," should be time, "2," should be S, "3," should be I, "4," R
#  print(head(incidence_modelled))
#  lambda <- incidence_modelled +
#    rexp(n = length(incidence_modelled), rate = exp_noise)
#  dpois(x = observed$Infected, lambda = lambda, log = TRUE)
#}
```




Inferring parameters

Using these pieces, we can set up a particle filter as follows:

```{r}
n_particles <- 1000
filter <- mcstate::particle_filter$new(data = sir_data,
                                       model = gen_sir,
                                       n_particles = n_particles,
                                       compare = case_compare,
                                       seed = 1L)
```

We can now run the particle filter forward, which will run and resample 100 trajectories, and return the final likelihood. It is important to set the correct time-step `dt` here, as we are using 0.25, rather than the default value of 1 defined in the odin model. We will also save the history, which allows us to plot the particle trajectories, as well as use the likelihood:

```{r}
filter$run(save_history = TRUE, pars = list(dt = dt))
```
If we plot these along with the data, we can see compared to above that only trajectories consistent with the data are kept, and the variance between particles has been reduced compared to the simulations shown in the `odin.dust` vignette:

```{r}
plot_particle_filter <- function(history, true_history, times, obs_end = NULL) {
  if (is.null(obs_end)) {
    obs_end <- max(times)
  }
  par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
  cols <- c(S = "#8c8cd9", I = "#cc0044", R = "#999966", TH = "#000000")
  matplot(times, t(history[2, , -1]), type = "l",
          xlab = "Time", ylab = "Number of individuals",
          col = cols[["S"]], lty = 1, ylim = range(history))
  matlines(times, t(history[3, , -1]), col = cols[["I"]], lty = 1)
  matlines(times, t(history[4, , -1]), col = cols[["R"]], lty = 1)
  matpoints(times[1:obs_end], (true_history), pch = 19, col = cols[["TH"]])
  legend("left", lwd = 1, col = cols, legend = names(cols), bty = "n")
}
#true_history <- readRDS("seir_true_history.rds")
true_history <- df_SIRinfected$Infected
plot_particle_filter(filter$history(), true_history, df_infected$day)

```

```{r}
filter$run(save_history = TRUE, pars = list(dt = dt, beta = 0.15, gamma = 0.1))

plot_particle_filter(filter$history(), true_history, df_infected$day)

```


## MCMC for fitting

```{r}
beta <- mcstate::pmcmc_parameter("beta", 0.2, min = 0.1)
gamma <- mcstate::pmcmc_parameter("gamma", 0.1, min = 0.075, prior = function(p)
  dgamma(p, shape = 1, scale = 0.2, log = TRUE))
#gamma <- mcstate::pmcmc_parameter("gamma", 0.1, min = 0.1)


proposal_matrix <- diag(0.1, 2)
mcmc_pars <- mcstate::pmcmc_parameters$new(list(beta = beta, gamma = gamma),
                                           proposal_matrix)
mcmc_pars$initial()
mcmc_pars$model(mcmc_pars$initial())
```


```{r}
n_steps <- 500
n_burnin <- 200
```

```{r}
control <- mcstate::pmcmc_control(
    n_steps,
    save_state = TRUE,
    save_trajectories = TRUE,
    progress = TRUE)
pmcmc_run <- mcstate::pmcmc(mcmc_pars, filter, control = control)
#> Running chain 1 / 1
#> Finished 500 steps in 14 secs
plot_particle_filter(pmcmc_run$trajectories$state, true_history, df_infected$day)
```

#it's a fit! :-)

# plot beta over gamma

```{r}
plot(pmcmc_run$pars[,"beta"],pmcmc_run$pars[,"gamma"])
print(length(pmcmc_run$pars[,"beta"]))
print(unique(pmcmc_run$pars[,"beta"]))
```



```{r}
processed_chains <- mcstate::pmcmc_thin(pmcmc_run, burnin = n_burnin, thin = 2)
parameter_mean_hpd <- apply(processed_chains$pars, 2, mean)
parameter_mean_hpd
```
# hm, quite different from the values of the original model (beta=0.2, gamma=0.1)... hm.

```{r}
library(coda)
mcmc1 <- coda::as.mcmc(cbind(pmcmc_run$probabilities, pmcmc_run$pars))

summary(mcmc1)
```

```{r}
plot(mcmc1)
```
# Tuning PMCMC

```{r}
print("effective Sample Size")
coda::effectiveSize(mcmc1)
print("Rejection Rate")
1 - coda::rejectionRate(mcmc1)
```
# the sample size of 4-7 (for beta and gamma) are way too small. We'd want something like 100.
#the acceptance rate is way too low (not sure which value exactly, prob esp the ones for beta and gamma), they should be at 5% or so

```{r}
proposal_matrix <- cov(pmcmc_run$pars)
mcmc_pars <- mcstate::pmcmc_parameters$new(
  list(beta = beta, gamma = gamma),
  proposal_matrix)
proposal_matrix
```

```{r}
control <- mcstate::pmcmc_control(
    n_steps,
    save_state = TRUE,
    save_trajectories = TRUE,
    progress = TRUE,
    n_chains = 4)
pmcmc_tuned_run <- mcstate::pmcmc(mcmc_pars, filter, control = control)


mcmc2 <- coda::as.mcmc(cbind(
  pmcmc_tuned_run$probabilities, pmcmc_tuned_run$pars))

summary(mcmc2)
```

```{r}
plot(mcmc2)
```

```{r}
print("effective Sample Size")
coda::effectiveSize(mcmc2)
print("Rejection Rate")
1 - coda::rejectionRate(mcmc2)
```
# plot beta over gamma

```{r}
plot(pmcmc_tuned_run$pars[,"beta"],pmcmc_tuned_run$pars[,"gamma"])
print(length(pmcmc_tuned_run$pars[,"beta"]))
```



# checking the fit of the tuned run visually:

```{r}
plot_particle_filter(pmcmc_tuned_run$trajectories$state, true_history, df_infected$day)
```











# well, this is not great either. hm.
# something is wrong with gamma. I don't really know what. Investigate please!

# 15.02.23: fit is much better! (the dt in the sir model definition (sir.R) of the transition prob was maybe wrong...?)
# the fit is okay but the value for beta is really off (0.4 instead of 0.15). That is weird because gamma=1 is correct. A time step thing? hmm

# update(time) <- (step + 1) * dt and update(time) <- (step + 1) are not the same, even though dt=1
# ....?!


### plot the pandemic with the beta and gamma values that the fit produced
### hahaha, I think I just tricked myself :D
### so, the beta was correctly fitted. I just thought that it was a different value, but it was, formerly, actually set to 4

### next step would be to fit to incidence instead of number of cases
# because this is the information we usually have available in a pandemic/endemic




```{r fig.height=5, fig.width=7}
n_times <- 200
x <- array(NA, dim = c(sir_model$info()$len, n_particles, n_times))

for (t in seq_len(n_times)) {
  x[ , , t] <- sir_model$run(t)
}
time <- x[1, 1, ]
x <- x[-1, , ]

par(mar = c(4.1, 5.1, 0.5, 0.5), las = 1)
cols <- c(S = "#8c8cd9", I = "#cc0044", R = "#999966")
matplot(time, t(x[1, , ]), type = "l",
         xlab = "Time", ylab = "Number of individuals",
         col = cols[["S"]], lty = 1, ylim = range(x))
matlines(time, t(x[2, , ]), col = cols[["I"]], lty = 1)
matlines(time, t(x[3, , ]), col = cols[["R"]], lty = 1)
legend("left", lwd = 1, col = cols, legend = names(cols), bty = "n")
```
























### effective sample size and run time for different number of particles

```{r}
npart <- c(16, 32, 64, 128)
MCMC_runs <- c(NA,NA,NA,NA)

for (i in 1:length(npart)) {
 n_particles <- npart[i]
 n_steps <- 20000 / n_particles
 n_burnin <- 1/5 * n_steps
 pmcmc_npart_run <- mcstate::pmcmc(mcmc_pars, filter, control = control) 
 MCMC_runs[i] <- coda::as.mcmc(cbind(pmcmc_npart_run$probabilities, pmcmc_npart_run$pars)) # prob does not work like that! 
 # please re-think what you want to do and why :-) 
}
```

```{r}
coda::effectiveSize(MCMC_runs[1])
```


