---
title: "Navajo Data Processing"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
path_to_data <- "~/Documents/PhD_Project/Data/"
```

# read in meta data (esp. serotype, sequencing year)
```{r}
library(readxl)
Navajo_meta_data <- read_excel(paste(path_to_data, "StrepPneumo_Navajo/metadata.xlsx", sep = ""))
Navajo_time_points <- sort(unique(Navajo_meta_data$Year))

Navajo_SeqYear_dict <- Navajo_meta_data$Year
names(Navajo_SeqYear_dict) <- Navajo_meta_data$Accession

Navajo_meta_IDtoInd <- 1:nrow(Navajo_meta_data)
names(Navajo_meta_IDtoInd) <- Navajo_meta_data$Accession
```


```{r}
Navajo_time_seq <- Navajo_time_points[1]:Navajo_time_points[length(Navajo_time_points)]
seq_no_per_year <- rep(0, length(Navajo_time_seq))
names(seq_no_per_year) <- Navajo_time_seq
for (i in 1:length(Navajo_time_seq)) {
  seq_no_per_year[i] <- length(which(Navajo_meta_data$Year==Navajo_time_seq[i]))
}
barplot(seq_no_per_year, xlab = "Sequencing Years", ylab = "Sequence Counts", col = rainbow(15))
# PCV7 was introduced in 2000
# PCV13 was introduced in 2010
# according to Azarian et al. (2018) https://doi.org/10.1371/journal.ppat.1006966
# 
Navajo_time_points_selected <- Navajo_time_seq
Navajo_time_points_preVac1 <- Navajo_time_seq[1:4]
Navajo_time_points_preVac2 <- Navajo_time_seq[1:13]
```


# read in PopPUNK results
```{r}
PP_all <- read.csv(paste(path_to_data, "StrepPneumo_NavajoNew/PopPUNK/poppunk_clusters_clusters.csv",sep = ""))
PP_taxon_ind_dict <- 1:nrow(PP_all)
names(PP_taxon_ind_dict) <- PP_all$Taxon

PP_is_Navajo_dict <- rep(FALSE,nrow(PP_all))
names(PP_is_Navajo_dict) <- PP_all$Taxon

for (i in 1:length(Navajo_meta_data$Accession)) {
  PP_is_Navajo_dict[Navajo_meta_data$Accession[i]] <- TRUE
}
Navajo_PP <- PP_all[PP_is_Navajo_dict,]
no_Navajo_PP <- length(unique(Navajo_PP$Cluster)) 
# 73 clusters

Navajo_seq_clusters_dict <- Navajo_PP$Cluster
names(Navajo_seq_clusters_dict) <- Navajo_PP$Taxon
```

```{r}
saveRDS(Navajo_meta_data, "Navajo_meta_data.rds")
```

```{r}
# Add Serotype Information
Navajo_Serotype_dict <- Navajo_meta_data$Sero
names(Navajo_Serotype_dict) <- Navajo_meta_data$Accession

Navajo_PP$Serotype <- Navajo_Serotype_dict[Navajo_PP$Taxon]

# Add VT information
PCV7_VTs_1 <- rep("NVT",length(unique(Navajo_meta_data$Sero)))
names(PCV7_VTs_1) <- unique(Navajo_meta_data$Sero)
PCV7_VTs_1[c("4", "6A","6B", "9V", "14", "18C", "19F", "23F")] <- "VT"

#Navajo_VT_PCV7_dict <- Navajo_meta_data$PCV7.actual
#names(Navajo_VT_PCV7_dict) <- Navajo_meta_data$Accession

#Navajo_PP$VT <- Navajo_VT_PCV7_dict[Navajo_PP$Taxon]
Navajo_PP$VT <- PCV7_VTs_1[Navajo_Serotype_dict[Navajo_PP$Taxon]]

# Add PCV13 info
PCV13_VTs_1 <- rep("NVT",length(unique(Navajo_meta_data$Sero)))
names(PCV13_VTs_1) <- unique(Navajo_meta_data$Sero)
PCV13_VTs_1[c("1", "3", "4","5","6A","6B", "7F", "9V", "14", "18C", "19A", "19F", "23F")] <- "VT"

Navajo_PP$VT_PCV13 <- PCV13_VTs_1[Navajo_Serotype_dict[Navajo_PP$Taxon]]

# Add sequencing year info
Navajo_Year_dict <- Navajo_meta_data$Year
names(Navajo_Year_dict) <- Navajo_meta_data$Accession

Navajo_PP$SeqYear <- Navajo_Year_dict[Navajo_PP$Taxon]

saveRDS(Navajo_PP,"Navajo_PP.rds")
```


# read in ggCaller results
```{r}
Navajo_ggCaller <- read.csv(paste(path_to_data, "StrepPneumo_NavajoNew/ggCaller_output/gene_presence_absence.csv", sep = ""), header=FALSE)
```

```{r}
# converting the gene presence absence matrix into a boolean df (0 = gene not present, 1 = gene present)
convert_to_bool <- function(x){
  if (x=="") 0 else 1
}

Navajo_ggCaller_bool <- Navajo_ggCaller[,c(-2,-3)]
Navajo_ggCaller_bool[-1,-1] <- apply(Navajo_ggCaller_bool[-1,-1],c(1,2), convert_to_bool)
#Navajo_ggCaller_bool[1,-1] <- unlist(strsplit(Navajo_ggCaller_bool[1,-1],".contigs_velvet"))

Navajo_ggCaller_bool[1,-1] <- sapply(Navajo_ggCaller_bool[1,-1], strsplit, split = ".contigs_velvet") # remove contigs appendix
colnames(Navajo_ggCaller_bool) <- Navajo_ggCaller_bool[1,]

PP_taxon_in_ggC <- rep(FALSE, length(Navajo_PP$Taxon))
names(PP_taxon_in_ggC) <- Navajo_PP$Taxon

#> length(PP_taxon_in_ggC)
#[1] 937

for (i in 2:ncol(Navajo_ggCaller_bool)) {
  PP_taxon_in_ggC[colnames(Navajo_ggCaller_bool)[i]] <- TRUE
}

#> length(PP_taxon_in_ggC)
#[1] 951

# why??

# hm, maybe filter by accession numbers that I actually had?
# running PP in that --include_database mode really seems to risk that I include stuff that wasn't actually in Navajo...

ggC_has_meta <- rep(FALSE, (ncol(Navajo_ggCaller_bool)-1))
names(ggC_has_meta) <- colnames(Navajo_ggCaller_bool)[-1]
for (i in 1:nrow(Navajo_meta_data)) {
  ggC_has_meta[Navajo_meta_data$Accession[i]] <- TRUE
}

# there are 14 sequences in ggCaller for which I do not have any meta data. why?? 
# Well, okay, I only have metadate for 937 sequences. So I am only going to consider those.

meta_has_ggC <- rep(FALSE, nrow(Navajo_meta_data))
names(meta_has_ggC) <- Navajo_meta_data$Accession
for (i in 2:ncol(Navajo_ggCaller_bool)) {
  meta_has_ggC[colnames(Navajo_ggCaller_bool)[i]] <- TRUE
}

which(!meta_has_ggC) # no more left. All sequences that have meta data are in ggC

which(!ggC_has_meta) # there are still some sequences for which I don't have meta data. but well.
# filter ggCaller results by the ones 
Navajo_ggCallerPP_bool <- data.frame(matrix(0, nrow = nrow(Navajo_ggCaller_bool), ncol = length(Navajo_PP$Taxon)+1))
Navajo_ggCallerPP_bool[-1,-1] <- Navajo_ggCaller_bool[-1,(which(ggC_has_meta))]
colnames(Navajo_ggCallerPP_bool) <- c("Gene",Navajo_PP$Taxon)
Navajo_ggCallerPP_bool[1,] <- c("Gene",Navajo_PP$Taxon)
Navajo_ggCallerPP_bool[,1] <- Navajo_ggCaller_bool[,1]
```

```{r}
saveRDS(Navajo_ggCallerPP_bool, "Navajo_ggCaller_bool.rds")
```

# split ggCaller gene presence absence matrix into different time points
```{r}
Navajo_ggCaller_byYear <- vector(mode = "list", length = length(Navajo_time_points_selected))

for (i in 1:length(Navajo_time_points_selected)) {
  Navajo_ggCaller_year <- data.frame(matrix(0, nrow = nrow(Navajo_ggCallerPP_bool), ncol = length(which(Navajo_meta_data$Year==Navajo_time_points_selected[i]))+1)) # initialize data frame
  Navajo_ggCaller_year[1,-1] <- Navajo_ggCallerPP_bool[1,c(FALSE,Navajo_SeqYear_dict[unlist(Navajo_ggCallerPP_bool[1,-1])]==Navajo_time_points_selected[i])] # fill in first row (seq names)
  Navajo_ggCaller_year[-1,1] <- Navajo_ggCallerPP_bool[-1,1] # fill in first column (gene cluster names)
  Navajo_ggCaller_year[-1,-1] <- Navajo_ggCallerPP_bool[-1,c(FALSE,Navajo_SeqYear_dict[unlist(Navajo_ggCallerPP_bool[1,-1])]==Navajo_time_points_selected[i])] # fill in presence absence information
  Navajo_ggCaller_byYear[[i]] <- Navajo_ggCaller_year
}
```

# find intermediate frequency genes
# I will combine all pre-vaccine data points for this
```{r}
sum_as_int <- function(x){
  sum(as.integer(x))
}

Navajo_gene_freq_preVac <- rep(0, nrow(Navajo_ggCaller_byYear[[1]])-1)
preVac_datapoints <- 0
for (i in 1:length(Navajo_time_points_preVac1)) {
  Navajo_gene_freq_preVac <- Navajo_gene_freq_preVac + apply(Navajo_ggCaller_byYear[[i]][-1,-1],1, sum_as_int)
  preVac_datapoints <- preVac_datapoints + length(Navajo_ggCaller_byYear[[i]][1,-1])
}
Navajo_gene_freq_preVac <- Navajo_gene_freq_preVac / preVac_datapoints

Navajo_gene_filter <- as.integer(Navajo_gene_freq_preVac<=0.95 & Navajo_gene_freq_preVac>=0.05)
# 1689 intermediate-frequency genes (of 5713)
```

# create intermediate frequency dataframes
```{r}
Navajo_ggCaller_byYear_intermed <- vector(mode = "list", length = length(Navajo_time_points_selected))

for (i in 1:length(Navajo_time_points_selected)) {
  Navajo_ggCaller_year <- data.frame(matrix(0, nrow = sum(Navajo_gene_filter)+1,ncol = length(which(Navajo_meta_data$Year==Navajo_time_points_selected[i]))+1))
  if(ncol(Navajo_ggCaller_byYear[[i]])>1){
    Navajo_ggCaller_year <-  Navajo_ggCaller_byYear[[i]][c(1,which(Navajo_gene_filter==1)+1),]
  }
  else{
    Navajo_ggCaller_year <-  data.frame(matrix(Navajo_ggCaller_byYear[[6]][c(1,which(Navajo_gene_filter==1)+1),1], nrow = sum(Navajo_gene_filter)+1,ncol = 1))
  }
  Navajo_ggCaller_byYear_intermed[[i]] <- Navajo_ggCaller_year 
}

Navajo_ggCaller_intermed <- data.frame(matrix(0, nrow = sum(Navajo_gene_filter)+1,ncol = ncol(Navajo_ggCallerPP_bool)))
Navajo_ggCaller_intermed <-  Navajo_ggCallerPP_bool[c(1,which(Navajo_gene_filter==1)+1),]

# create pre_Vac intermed frequency data frame
Navajo_ggCaller_intermed_preVac <- data.frame(matrix(0, nrow = sum(Navajo_gene_filter)+1,ncol = (sum(unlist(lapply(Navajo_ggCaller_byYear_intermed,ncol))[1:length(Navajo_time_points_preVac1)]) - length(Navajo_time_points_preVac1) + 1)))
col_count <- 2
for (i in 1:length(Navajo_time_points_preVac1)) {
  if(ncol(Navajo_ggCaller_byYear_intermed[[i]])>1){
    Navajo_ggCaller_intermed_preVac[,(col_count:(col_count + ncol(Navajo_ggCaller_byYear_intermed[[i]])-1))] <- Navajo_ggCaller_byYear_intermed[[i]][,-1]
    col_count <- col_count + ncol(Navajo_ggCaller_byYear_intermed[[i]])
  }
}
Navajo_ggCaller_intermed_preVac[,1] <-  Navajo_ggCaller_byYear_intermed[[1]][,1]

saveRDS(Navajo_ggCaller_intermed, "Navajo_ggCaller_intermed.rds")
```

# create consensus gene presence absence matrices for the clusters
```{r}
cons_genomes <- function(x){
  as.double(ceiling(median(as.integer(x))))
}

# overall
Navajo_ggCaller_intermed_consensus <- data.frame(matrix(0, nrow = sum(Navajo_gene_filter)+1, ncol = no_Navajo_PP+1))
Navajo_ggCaller_intermed_consensus[1,-1] <- paste("SeqCl_",unique(Navajo_PP$Cluster),sep = "")

for (i in unique(Navajo_PP$Cluster)) {
  Navajo_ggCaller_intermed_consensus[-1,(which(unique(Navajo_PP$Cluster) == i))+1] <- apply(as.matrix(Navajo_ggCaller_intermed[-1,c(FALSE,Navajo_seq_clusters_dict[unlist(Navajo_ggCaller_intermed[1,-1])]==i)]), 1, cons_genomes)

}
Navajo_ggCaller_intermed_consensus[,1] <- Navajo_ggCaller_intermed[,1]

#by year
Navajo_ggCaller_byYear_intermed_consensus <- vector(mode = "list", length = length(Navajo_time_points_selected))

for (i in 1:length(Navajo_time_points_selected)) {
  Navajo_ggCaller_year <- data.frame(matrix(0, nrow = sum(Navajo_gene_filter)+1,ncol = no_Navajo_PP+1))
  Navajo_ggCaller_year[1,-1] <- paste("SeqCl_",unique(Navajo_PP$Cluster),sep = "")
  for (j in unique(Navajo_PP$Cluster)) {
    Navajo_ggCaller_year[-1,(which(unique(Navajo_PP$Cluster) == j))+1] <- apply(as.matrix(Navajo_ggCaller_byYear_intermed[[i]][-1,c(FALSE,Navajo_seq_clusters_dict[unlist(Navajo_ggCaller_byYear_intermed[[i]][1,-1])]==j)]), 1, cons_genomes)
  }
  Navajo_ggCaller_year[,1] <- Navajo_ggCaller_byYear_intermed[[i]][,1]
  Navajo_ggCaller_byYear_intermed_consensus[[i]] <- Navajo_ggCaller_year
}

saveRDS(Navajo_ggCaller_intermed_consensus,"Navajo_ggCaller_intermed_consensus.rds")

# intermed gene name list for mapping ggCaller
Navajo_ggC_intermed_gene_names <- Navajo_ggCaller_intermed_consensus[-1,1]
saveRDS(Navajo_ggC_intermed_gene_names, file = "Navajo_ggC_intermed_gene_names.rds")
```

# Calculate VT/NVT for clusters
```{r}
# create dictionary for VTs of the PCV10
# 1, 4, 5, 6b, 7f, 9V, 14, 18c, 19f and 23f serotypes (source: NCBI)
# PCV13: (1, 3, 4, 5, 6A and 6B, 7F, 9V, 14, 19A, and 19F, 18C, and 23F) (source: https://www.ncbi.nlm.nih.gov/books/NBK507794/)
#PCV7: 4, 6B, 9V, 14, 18C, 19F and 23F. (source: https://journals.plos.org/plospathogens/article?id=10.1371/journal.ppat.1006966#sec018)
PCV13_VTs <- rep(0,length(unique(Navajo_meta_data$Sero)))
names(PCV13_VTs) <- unique(Navajo_meta_data$Sero)
PCV13_VTs[c("1", "3", "4","5","6A","6B", "7F", "9V", "14", "18C", "19A", "19F", "23F")] <- 1

#PCV7_VTs <- rep(0,length(unique(Navajo_meta_data$Sero)))
#names(PCV7_VTs) <- unique(Navajo_meta_data$Sero)
#PCV7_VTs[c("4","6B", "9V", "14", "18C", "19F", "23F")] <- 1

# add 6A to PCV7 because there is strong cross-immunity btw PVC7 and 6A (4.Croucher, N. J. et al. Population genomics of post-vaccine changes in pneumococcal epidemiology. Nat. Genet. 45, 656–663 (2013).)
PCV7_VTs <- rep(0,length(unique(Navajo_meta_data$Sero)))
names(PCV7_VTs) <- unique(Navajo_meta_data$Sero)
PCV7_VTs[c("4", "6A","6B", "9V", "14", "18C", "19F", "23F")] <- 1

Navajo_SeqToSero_dict <- Navajo_meta_data$Sero
names(Navajo_SeqToSero_dict) <- Navajo_meta_data$Accession

# calculate Vaccine Type consensus for clusters
Navajo_VT <- rep(0, no_Navajo_PP)

#for (i in unique(Navajo_PP$Cluster)){
#  Navajo_VT[(which(unique(Navajo_PP$Cluster) == i))] <- ceiling(median(as.integer(PCV10_VTs[Navajo_SeqToSero_dict[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]]])))
#}
# I apparently did this for PCV10. That doesn't make much sense
for (i in unique(Navajo_PP$Cluster)){
  Navajo_VT[(which(unique(Navajo_PP$Cluster) == i))] <- ceiling(median(as.integer(PCV7_VTs[Navajo_SeqToSero_dict[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]]])))
}

Navajo_VT2 <- rep(0, no_Navajo_PP)
for (i in unique(Navajo_PP$Cluster)){
  Navajo_VT2[(which(unique(Navajo_PP$Cluster) == i))] <- ceiling(median(as.integer(PCV13_VTs[Navajo_SeqToSero_dict[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]]])))
}
# 1 means VT
# 0 means NVT
# if cluster is 50/50 it will count as VT
```

```{r}
#save VTs
saveRDS(Navajo_VT, file = "Navajo_VT.rds")
saveRDS(Navajo_VT2, file = "Navajo_VT2.rds")
```

# calculate cluster frequencies for different years
```{r}
Navajo_cluster_freqs <- vector(mode = "list", length = length(Navajo_time_points_selected))

for(j in 1:length(Navajo_time_points_selected)){
  Navajo_cluster_freqs[[j]] <- rep(NA, no_Navajo_PP)
  for (i in unique(Navajo_PP$Cluster)){
    Navajo_cluster_freqs[[j]][which(unique(Navajo_PP$Cluster) == i)] <- length(which(Navajo_meta_data[Navajo_meta_IDtoInd[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]],]$Year==Navajo_time_points_selected[j]))
  }
}

for(j in 1:length(Navajo_time_points_selected)){
  if(sum(Navajo_cluster_freqs[[j]])==0){
    Navajo_cluster_freqs[[j]] <- rep(NA, no_Navajo_PP)
  }
}

#Navajo_cluster_freqs[[6]] <- rep(NA, no_Navajo_PP)
# 901 data points in total
Navajo_cluster_freqs_preVac <- rep(0, no_Navajo_PP)
for (i in 1:length(Navajo_time_points_preVac1)) {
  Navajo_cluster_freqs_preVac <- Navajo_cluster_freqs_preVac + replace(Navajo_cluster_freqs[[i]],which(is.na(Navajo_cluster_freqs[[i]])),0)
}

for (i in 1:length(Navajo_time_points_selected)) {
  file_name <- paste("Navajo_cluster_freqs_", as.character(i),".rds", sep = "")
  saveRDS(Navajo_cluster_freqs[[i]], file_name)
}


# repeat this but split VTs from NVTs per cluster
Navajo_VT_cluster_freqs <- vector(mode = "list", length = length(Navajo_time_points_selected))
Navajo_NVT_cluster_freqs <- vector(mode = "list", length = length(Navajo_time_points_selected))
for(j in 1:length(Navajo_time_points_selected)){
  Navajo_VT_cluster_freqs[[j]] <- rep(NA, no_Navajo_PP)
  Navajo_NVT_cluster_freqs[[j]] <- rep(NA, no_Navajo_PP)
  for (i in unique(Navajo_PP$Cluster)){
    Navajo_VT_cluster_freqs[[j]][which(unique(Navajo_PP$Cluster) == i)] <- length(intersect(which(Navajo_meta_data[Navajo_meta_IDtoInd[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]],]$Year==Navajo_time_points_selected[j]),  which(PCV7_VTs[Navajo_meta_data[Navajo_meta_IDtoInd[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]],]$Sero]==1)))
  Navajo_NVT_cluster_freqs[[j]][which(unique(Navajo_PP$Cluster) == i)] <- length(intersect(which(Navajo_meta_data[Navajo_meta_IDtoInd[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]],]$Year==Navajo_time_points_selected[j]),  which(PCV7_VTs[Navajo_meta_data[Navajo_meta_IDtoInd[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]],]$Sero]==0)))
  }
}

for(j in 1:length(Navajo_time_points_selected)){
  if(sum(Navajo_VT_cluster_freqs[[j]])==0){
    Navajo_VT_cluster_freqs[[j]] <- rep(NA, no_Navajo_PP)
  }
}

for(j in 1:length(Navajo_time_points_selected)){
  if(sum(Navajo_NVT_cluster_freqs[[j]])==0){
    Navajo_NVT_cluster_freqs[[j]] <- rep(NA, no_Navajo_PP)
  }
}

for (i in 1:length(Navajo_time_points_selected)) {
  file_name <- paste("Navajo_VT_cluster_freqs_", as.character(i),".rds", sep = "")
  saveRDS(Navajo_VT_cluster_freqs[[i]], file_name)
  
  file_name <- paste("Navajo_NVT_cluster_freqs_", as.character(i),".rds", sep = "")
  saveRDS(Navajo_NVT_cluster_freqs[[i]], file_name)
}



# do the same for Strain x Serotype model
Navajo_cluster_freqs_sero <- vector(mode = "list", length = length(Navajo_time_points_selected))

for(j in 1:length(Navajo_time_points_selected)){
  Navajo_cluster_freqs_sero[[j]] <- matrix(NA, ncol =  length(unique(Navajo_PP$Serotype)), nrow = no_Navajo_PP)
  for (i in unique(Navajo_PP$Cluster)){
    for (k in unique(Navajo_PP$Serotype)) {
      Navajo_cluster_freqs_sero[[j]][which(unique(Navajo_PP$Cluster) == i),which(unique(Navajo_PP$Serotype) == k)] <- length(which(Navajo_meta_data[Navajo_meta_IDtoInd[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]],][Navajo_meta_data[Navajo_meta_IDtoInd[Navajo_PP[Navajo_PP$Cluster == i,"Taxon"]],]$Sero==k,]$Year==Navajo_time_points_selected[j]))
    }
  }
}
for(j in 1:length(Navajo_time_points_selected)){
  if(sum(Navajo_cluster_freqs_sero[[j]])==0){
    Navajo_cluster_freqs_sero[[j]] <- matrix(NA, ncol =  length(unique(Navajo_PP$Serotype)), nrow = no_Navajo_PP)
  }
}

for (i in 1:length(Navajo_time_points_selected)) {
  file_name <- paste("Navajo_cluster_freqs_sero_", as.character(i),".rds", sep = "")
  saveRDS(Navajo_cluster_freqs_sero[[i]], file_name)
}


```

#create initial population
```{r}
### create initial population that is based on the first time point data set
# sample from it with an Poisson process
expand_factor <- 15000 / sum(Navajo_cluster_freqs_preVac)
exp_noise <- 10
Navajo_model_start_pop <- (sapply((Navajo_cluster_freqs_preVac + rexp(n = length(Navajo_cluster_freqs_preVac), rate = exp_noise)) * expand_factor, rpois, n=1))

# visual check for the "sampling"
plot(Navajo_cluster_freqs_preVac/sum(Navajo_cluster_freqs_preVac))
points(Navajo_model_start_pop/sum(Navajo_model_start_pop), col = "red")

expand_factor <- 15000 / sum(Navajo_cluster_freqs[[1]])
exp_noise <- 10
Navajo_model_start_pop <- (sapply((Navajo_cluster_freqs[[1]] + rexp(n = length(Navajo_cluster_freqs[[1]]), rate = exp_noise)) * expand_factor, rpois, n=1))

# visual check for the "sampling"
plot(Navajo_cluster_freqs[[1]]/sum(Navajo_cluster_freqs[[1]]))
points(Navajo_model_start_pop/sum(Navajo_model_start_pop), col = "red")

# I will only use the first time point for initialising the population. But I need to go back and change that later?

saveRDS(Navajo_model_start_pop, "Navajo_model_start_pop.rds")
```


# calculate delta ranking
# I will do last data point - all pre_vac data points combined
```{r}
# For strong and weak selection in the model, I need to calculate the beta statistics.

# calculate gene frequencies first, separate for three time points
Navajo_gene_freq_preVac <- apply(Navajo_ggCaller_intermed_preVac[-1,-1], 1, sum_as_int)
Navajo_gene_freq_postVac <- apply(Navajo_ggCaller_byYear_intermed[[length(Navajo_ggCaller_byYear_intermed)]][-1,-1], 1, sum_as_int) + apply(Navajo_ggCaller_byYear_intermed[[length(Navajo_ggCaller_byYear_intermed)-1]][-1,-1], 1, sum_as_int) + apply(Navajo_ggCaller_byYear_intermed[[length(Navajo_ggCaller_byYear_intermed)-2]][-1,-1], 1, sum_as_int)


# first, calculate pre/peri and post vacc frequencies of genes:
Navajo_gene_freq_preVac_rel <- Navajo_gene_freq_preVac/sum(Navajo_gene_freq_preVac)
Navajo_gene_freq_postVac_rel <- Navajo_gene_freq_postVac/sum(Navajo_gene_freq_postVac)

# calculate delta statistic (refer to Corander et al. for more info)
Navajo_delta_data <- (Navajo_gene_freq_postVac_rel - Navajo_gene_freq_preVac_rel) ^ 2 / (1 - Navajo_gene_freq_preVac_rel * (1 - Navajo_gene_freq_preVac_rel))
Navajo_delta_ranking <- rank(Navajo_delta_data)

saveRDS(Navajo_delta_ranking,"Navajo_delta_ranking.rds")
```

### Data processing for PPxSero model
```{r}
PopPUNK_clusters <- readRDS("Navajo_PP.rds")
Navajo_model_start_pop <- readRDS("Navajo_model_start_pop.rds")
mass_clusters <- length(unique(PopPUNK_clusters$Cluster))
Navajo_avg_cluster_freq <- rep(1/mass_clusters, mass_clusters)


PPxSero_matrix <- matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))
# sort(unique(PopPUNK_clusters$Serotype))
# [1] "10A"   "11A"   "14"    "15A"   "15B/C" "15F"   "16F"   "17F"   "18C"   "19A"   "19F"   "21"    "22F"   "23A"
#[15] "23B"   "23F"   "3"     "31"    "33F"   "34"    "35B"   "35F"   "37"    "38"    "6A"    "6B"    "6C"    "7C" 
#[29] "7F"    "9N"    "9V"    "NT"

SerotypeToIndex_dict <- 1:length(unique(PopPUNK_clusters$Serotype))
names(SerotypeToIndex_dict) <- (unique(PopPUNK_clusters$Serotype))

IndexToSerotype_dict <- (unique(PopPUNK_clusters$Serotype))
names(IndexToSerotype_dict) <- 1:length(unique(PopPUNK_clusters$Serotype))

PPclustToIndex_dict <- 1:length(unique(PopPUNK_clusters$Cluster))
names(PPclustToIndex_dict) <- (unique(PopPUNK_clusters$Cluster))

for (i in 1:nrow(PopPUNK_clusters)) {
  PPxSero_matrix[PPclustToIndex_dict[as.character(PopPUNK_clusters$Cluster[i])],SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- 1
}
# matrix shows information whether PP and serotype co-occur in the dataset (1=TRUE)
# this can be used as the migration matrix:
#PPxSero_matrix_prob <- PPxSero_matrix / sum(PPxSero_matrix)
#PPxSero_matrix_prob <- matrix(sapply(PPxSero_matrix_prob,as.double), nrow = nrow(PPxSero_matrix), ncol = ncol(PPxSero_matrix))
#saveRDS(PPxSero_matrix_prob, "Navajo_PPsero_mig.rds")


PPxSero_matrix_prob2 <-  matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))
for (i in 1:length(unique(PopPUNK_clusters$Cluster))) {
  PPxSero_matrix_prob2[i,] <- ((PPxSero_matrix[i,] * Navajo_avg_cluster_freq[i])/sum(PPxSero_matrix[i,]))
}
saveRDS(PPxSero_matrix_prob2, "Navajo_PPsero_mig.rds")
# new migration matrix that is closer to original one

for (i in 1:nrow(PopPUNK_clusters)) {
  PPxSero_matrix[PPclustToIndex_dict[as.character(PopPUNK_clusters$Cluster[i])],SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- 1
}
# matrix shows information whether PP and serotype co-occur in the dataset (1=TRUE)
# this can be used as the migration matrix:
PPxSero_matrix_prob <- PPxSero_matrix / sum(PPxSero_matrix)
PPxSero_matrix_prob <- matrix(sapply(PPxSero_matrix_prob,as.double), nrow = nrow(PPxSero_matrix), ncol = ncol(PPxSero_matrix))
saveRDS(PPxSero_matrix_prob, "Navajo_PPsero_mig.rds")

# start population:
PPsero_startpop <- matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))

for (i in which(PopPUNK_clusters$SeqYear=="1998")) {
  PPsero_startpop[PPclustToIndex_dict[PopPUNK_clusters$Cluster[i]],SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- PPsero_startpop[PPclustToIndex_dict[PopPUNK_clusters$Cluster[i]],SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] + 1
}
#PPsero_startpop <- PPsero_startpop * 15000 / sum(PPsero_startpop)


PP_expand_factor <- 15000 / sum(PPsero_startpop)
exp_noise <- 10
#PPsero_startpop_corr <- matrix(sapply((PPsero_startpop + matrix(rexp(n = ncol(PPsero_startpop) * nrow(PPsero_startpop), rate = exp_noise), nrow= nrow(PPsero_startpop), ncol = ncol(PPsero_startpop))) * PP_expand_factor, rpois, n=1), nrow= nrow(PPsero_startpop), ncol = ncol(PPsero_startpop))
#PPsero_startpop_corr <- matrix(sapply(PPsero_startpop_corr,as.double), nrow= nrow(PPsero_startpop), ncol = ncol(PPsero_startpop))
#saveRDS(PPsero_startpop_corr, "PPsero_startpop.rds")

# this lead to a start population that is way too evenly distributed across the different clusters
# instead, I could maybe use the existing PP_model_start_pop and distribute it to the different serotypes?
PPsero_startpop2 <-  matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))
for (i in 1:length(unique(PopPUNK_clusters$Cluster))) {
  PPsero_startpop2[i,] <- round((PPxSero_matrix[i,] * Navajo_model_start_pop[i])/sum(PPxSero_matrix[i,]))
}
saveRDS(PPsero_startpop2, "Navajo_PPsero_startpop.rds")
# this is very close to the original PP_model_start_pop
# maybe I am giving out too much information because the serotype distribution is exactly the real one?
plot(rowSums(Navajo_cluster_freqs_sero[[1]])/sum(rowSums(Navajo_cluster_freqs_sero[[1]])))
points(rowSums(PPsero_startpop2)/sum(rowSums(PPsero_startpop2)), col = "red")

plot(colSums(Navajo_cluster_freqs_sero[[1]])/sum(colSums(Navajo_cluster_freqs_sero[[1]])))
points(colSums(PPsero_startpop2)/sum(colSums(PPsero_startpop2)), col = "red")
# the serotype distribution is quite far off

# same as PPsero_startpop6 for Mass
PPsero_startpop3 <- matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_expand_factor <- 15000 / sum(Navajo_cluster_freqs_sero[[1]])
exp_noise <- 1000
for (i in 1:nrow(PPsero_startpop3)) {
    PPsero_startpop3[i,] <- (sapply((Navajo_cluster_freqs_sero[[1]][i,] + rexp(n = length(Navajo_cluster_freqs_sero[[1]][i,]), rate = exp_noise)) * PP_expand_factor, rpois, n=1)) + ceiling(PPxSero_matrix_prob2[i,])
  #PPsero_startpop5[i,] <- PP_mass_cluster_freq_1_sero[i,] + ceiling(PPxSero_matrix_prob[i,])
}
saveRDS(PPsero_startpop3, "Navajo_PPsero_startpop.rds")

plot(rowSums(Navajo_cluster_freqs_sero[[1]])/sum(rowSums(Navajo_cluster_freqs_sero[[1]])))
points(rowSums(PPsero_startpop3)/sum(rowSums(PPsero_startpop3)), col = "red")

plot(colSums(Navajo_cluster_freqs_sero[[1]])/sum(colSums(Navajo_cluster_freqs_sero[[1]])))
points(colSums(PPsero_startpop3)/sum(colSums(PPsero_startpop3)), col = "red")


# serotype VT vector
#SeroVT <- rep(0, length(unique(PopPUNK_clusters$Serotype)))
#for (i in 1:nrow(PopPUNK_clusters)) {
#  SeroVT[SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- as.integer(PopPUNK_clusters$VT[i]=="VT")
#}
#saveRDS(SeroVT, "Navajo_SeroVT.rds")

# serotype VT vector
SeroVT <- rep(0, length(unique(PopPUNK_clusters$Serotype)))
for (i in 1:nrow(PopPUNK_clusters)) {
  SeroVT[SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- as.integer(PopPUNK_clusters$VT[i]=="VT")
}
saveRDS(SeroVT, "Navajo_SeroVT.rds")

# serotype VT vector PCV13
SeroVT_PCV13 <- rep(0, length(unique(PopPUNK_clusters$Serotype)))
for (i in 1:nrow(PopPUNK_clusters)) {
  SeroVT_PCV13[SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- as.integer(PopPUNK_clusters$VT_PCV13[i]=="VT")
}
saveRDS(SeroVT_PCV13, "Navajo_SeroVT_PCV13.rds")
```

```{r}
# Create PP clusters, split by serotype
#calculate the frequency of the gene clusters and year and serotype
Navajo_mass_cluster_freq_1_sero <- matrix(0,nrow = mass_clusters, ncol = length(unique(PopPUNK_clusters$Serotype)))

for (i in 1:mass_clusters){
  for (k in unique(PopPUNK_clusters$Serotype)){
  Navajo_mass_cluster_freq_1_sero[i,which(unique(PopPUNK_clusters$Serotype)==k)] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),][PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$Serotype ==k, ]$"SeqYear"==sort(unique(PopPUNK_clusters$SeqYear))[1]))
  }
}

#save sequence cluster frequencies
saveRDS(Navajo_mass_cluster_freq_1_sero, file = "Navajo_mass_cluster_freq_1_sero.rds")

PPxSero_matrix_prob2 <- readRDS("Navajo_PPsero_mig.rds")

Navajo_PPsero_startpop3 <- matrix(0,nrow = mass_clusters, ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_expand_factor <- 15000 / sum(Navajo_mass_cluster_freq_1_sero)
exp_noise <- 1000
for (i in 1:nrow(Navajo_PPsero_startpop3)) {
    Navajo_PPsero_startpop3[i,] <- (sapply((Navajo_mass_cluster_freq_1_sero[i,] + rexp(n = length(Navajo_mass_cluster_freq_1_sero[i,]), rate = exp_noise)) * PP_expand_factor, rpois, n=1)) + ceiling(PPxSero_matrix_prob2[i,])
}
plot(colSums(Navajo_mass_cluster_freq_1_sero)/sum(colSums(Navajo_mass_cluster_freq_1_sero)))
points(colSums(Navajo_PPsero_startpop3)/sum(colSums(Navajo_PPsero_startpop3)), col = "red")

plot(rowSums(Navajo_mass_cluster_freq_1_sero)/sum((Navajo_mass_cluster_freq_1_sero)))
points(rowSums(Navajo_PPsero_startpop3)/sum(rowSums(Navajo_PPsero_startpop3)), col = "red")

saveRDS(Navajo_PPsero_startpop3, file = "Navajo_PPsero_startpop3.rds")
```